---
title: "Assignment 10"
author: "Chee Kay Cheong"
output: github_document
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(caret)
library(glmnet)
```

# Part 1

### Data Exploratory Analyses

```{r}
# Load data file
load("./exposome.RData")

# Put all data frames into a list
df_list = list(covariates, exposome, phenotype) 

# Merge all data frames together
HW10 = df_list %>% 
  reduce(full_join, by = 'ID') %>% 
  select(-ID) %>% 
  mutate(hs_asthma = as_factor(hs_asthma))

# Check to see if there is missing values
miss.data = HW10[!complete.cases(HW10), ]
# no missing values in the dataset
```

The dataset `HW10` contains `r nrow(HW10)` observations and `r ncol(HW10)` variables.

```{r, results = 'hide'}
# Display rows and columns
dim(HW10)

# Summarize each variable in dataset
summary(HW10)
```

Descriptive measures:

* `h_trafload_preg_pow1over3` = Total traffic load of all roads in 100m buffer at pregnancy period (numeric)                                                                     
  + Min : 0.3458                                                                                                                                                                       
  + 1st Quartile: 33.6542                                                                                                                                                              
  + Median : 66.6101                                                                                                                                                                   
  + Mean : 75.5390                                                                                                                                                                     
  + 3rd Quartile: 113.0812                                                                                                                                                            
  + Max : 294.2705                                                                                                                                                                   

* `e3_asmokcigd_p_None` = maternal active tobacco smoke during pregnancy mean number of cigarette/day (numeric)                                                                       
  + Min : 0.000                                                                                                                                                                        
  + 1st Quartile: 0.000                                                                                                                                                                
  + Median : 0.000                                                                                                                                                                     
  + Mean : 0.494                                                                                                                                                                       
  + 3rd Quartile: 0.000                                                                                                                                                               
  + Max : 15.238

* `h_pm10_ratio_preg_None` = pm10 value (extrapolated back in time using ratio method) during pregnancy (numeric)                                                                     
  + Min : 8.066                                                                                                                                                                        
  + 1st Quartile: 17.535                                                                                                                                                               
  + Median : 23.018                                                                                                                                                                   
  + Mean : 23.504                                                                                                                                                                     
  + 3rd Quartile: 27.677                                                                                                                                                               
  + Max : 47.698

* `hs_pm10_wk_hs_h_None` = pm10 value (extrapolated back in time using ratio method) one week before hs test at home (numeric)
  + Min : 5.838                                                                                                                                                                  
  + 1st Quartile: 19.142                                                                                                                                                              
  + Median : 24.891                                                                                                                                                                   
  + Mean : 26.409                                                                                                                                                                     
  + 3rd Quartile: 32.131                                                                                                                                                            
  + Max : 211.297
  
* `h_lden_cat_preg_None` = Categorized lden (day, evening, night) at pregnancy period (numeric)
  + Min : 33.92                                                                                                                                                                       
  + 1st Quartile: 50.00                                                                                                                                                              
  + Median : 58.63                                                                                                                                                                   
  + Mean : 57.47                                                                                                                                                                     
  + 3rd Quartile: 64.36                                                                                                                                                            
  + Max : 77.40
  
* `hs_asthma` = Doctor diagnosed asthma (ever) (factor)
  + 0 (No) : 1159                                                                                                                                                                     
  + 1 (Yes) : 142

Examine correlation between variables:
```{r}
# Create correlation matrix
round(cor(HW10[c('h_trafload_preg_pow1over3', 'e3_asmokcigd_p_None', 'h_pm10_ratio_preg_None', 'hs_pm10_wk_hs_h_None', 'h_lden_cat_preg_None')]), 2)
```

# Part 2

### Research question 

What factors best predict the risk of asthma in children ages between 6-11 years of age from the HELIX cohort?

# Part 3

Since the `HW10` dataset contains more than 200 variables, I decided to perform feature selection using **LASSO** to reduce the number of variables in the dataset while retaining as much of the original variation as possible. This can help simplify further analysis (such as random forest, logistic regression, etc.) and improve computational efficiency.

### LASSO
```{r}
set.seed(123)

asthma_lasso = HW10

# Find correlated predictors to remove variables that are highly correlated
# Correlation can only be done with numeric variables, so need to filter out the non-numeric variables
asthma.numeric = 
  asthma_lasso %>% 
  select(where(is.numeric))

# Calculate correlations
correlations = cor(asthma.numeric, use = "complete.obs")

# Find any features that are correlated at 0.7 and above
high.correlations = findCorrelation(correlations, cutoff = 0.7) 

# Remove highly correlated features
asthma_lasso = asthma_lasso[ , -high.correlations]

# Partition data into 70/30 split
train.index = 
  asthma_lasso$hs_asthma %>% createDataPartition(p = 0.7, list = F)

train_data = asthma_lasso[train.index, ]
test_data = asthma_lasso[-train.index, ]
```

```{r}
set.seed(123)

control.settings = trainControl(method = "cv", number = 10, sampling = "up")

lambda = 10^seq(-3, -1, length = 100)

LASSO = train(hs_asthma ~ ., data = train_data, method = "glmnet", trControl = control.settings, preProcess = c("center", "scale"), tuneGrid = expand.grid(alpha = 1, lambda = lambda))

# Output best value of alpha & lambda
LASSO$finalModel$tuneValue
LASSO$bestTune
LASSO$results

confusionMatrix(LASSO)

coef(LASSO$finalModel, LASSO$bestTune$lambda)
varImp(LASSO)
```

```{r}
# Test model
test_outcome = predict(LASSO, test_data)

# Evaluation metric:
confusionMatrix(test_outcome, test_data$hs_asthma, positive = '1')
```

Based on the variable importance, we know which variables in the dataset are the most important to predict the outcome (asthma). From there, we can reduce the number of features in our dataset by selecting only the variables as indicated by the variable importance. Then, we can do further analysis using different algorithms such as random forest or logistic regression to predict our outcome of interest. 

