---
title: "Assignment 10"
author: "Chee Kay Cheong"
output: github_document
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(caret)
library(rpart.plot)
```

# Part 1

```{r}
# Load data file
load("./exposome.RData")

# Put all data frames into a list
df_list = list(covariates, exposome, phenotype) 

# Merge all data frames together
HW10 = df_list %>% 
  reduce(full_join, by = 'ID') %>% 
  select(-ID) %>% 
  mutate(hs_asthma = as_factor(hs_asthma))

# Check to see if there is missing values
miss.data = HW10[!complete.cases(HW10), ]
# no missing values in the dataset
```

The dataset `HW10` contains `r nrow(HW10)` observations and `r ncol(HW10)` variables.

```{r, results = 'hide'}
# Display rows and columns
dim(HW10)

# Summarize each variable in dataset
summary(HW10)
```

Descriptive measures:

* `h_trafload_preg_pow1over3` = Total traffic load of all roads in 100m buffer at pregnancy period (numeric)                                                                     
  + Min : 0.3458                                                                                                                                                                       
  + 1st Quartile: 33.6542                                                                                                                                                              
  + Median : 66.6101                                                                                                                                                                   
  + Mean : 75.5390                                                                                                                                                                     
  + 3rd Quartile: 113.0812                                                                                                                                                            
  + Max : 294.2705                                                                                                                                                                   

* `e3_asmokcigd_p_None` = maternal active tobacco smoke during pregnancy mean number of cigarette/day (numeric)                                                                       
  + Min : 0.000                                                                                                                                                                        
  + 1st Quartile: 0.000                                                                                                                                                                
  + Median : 0.000                                                                                                                                                                     
  + Mean : 0.494                                                                                                                                                                       
  + 3rd Quartile: 0.000                                                                                                                                                               
  + Max : 15.238

* `h_pm10_ratio_preg_None` = pm10 value (extrapolated back in time using ratio method) during pregnancy (numeric)                                                                     
  + Min : 8.066                                                                                                                                                                        
  + 1st Quartile: 17.535                                                                                                                                                               
  + Median : 23.018                                                                                                                                                                   
  + Mean : 23.504                                                                                                                                                                     
  + 3rd Quartile: 27.677                                                                                                                                                               
  + Max : 47.698

* `hs_pm10_wk_hs_h_None` = pm10 value (extrapolated back in time using ratio method) one week before hs test at home (numeric)
  + Min : 5.838                                                                                                                                                                  
  + 1st Quartile: 19.142                                                                                                                                                              
  + Median : 24.891                                                                                                                                                                   
  + Mean : 26.409                                                                                                                                                                     
  + 3rd Quartile: 32.131                                                                                                                                                            
  + Max : 211.297
  
* `h_lden_cat_preg_None` = Categorized lden (day, evening, night) at pregnancy period (numeric)
  + Min : 33.92                                                                                                                                                                       
  + 1st Quartile: 50.00                                                                                                                                                              
  + Median : 58.63                                                                                                                                                                   
  + Mean : 57.47                                                                                                                                                                     
  + 3rd Quartile: 64.36                                                                                                                                                            
  + Max : 77.40
  
* `hs_asthma` = Doctor diagnosed asthma (ever) (factor)
  + 0 (No) : 1159                                                                                                                                                                     
  + 1 (Yes) : 142

Examine correlation between variables:
```{r}
# Create correlation matrix
round(cor(HW10[c('h_trafload_preg_pow1over3', 'e3_asmokcigd_p_None', 'h_pm10_ratio_preg_None', 'hs_pm10_wk_hs_h_None', 'h_lden_cat_preg_None')]), 2)
```

# Part 2

Develop a single research question. It can be a prediction or hypothesis generating research question. It doesn't need to involve all features but it must involve at least one outcome from phenotype and 5 features from exposome (and covariates if relevant). This can be as simple or complex as a question as you want it to be. Clearly state the research question in a complete sentence outside of a code chunk, not as a comment within one.





